Crystals

### 2017-03-01 doc/ms.tex
Working on the document

\newcommand{\hypers}{\boldsymbol{\alpha}}
\newcommand{\params}{\boldsymbol{\theta}}
\newcommand{\data}{\boldsymbol{x}}
\newcommand{\info}{\boldsymbol{M}}
\newcommand{\x}{x}
\newcommand{\todo}{\color{orange} \bf}

@@

\section{The prior information}
In the Bayesian framework, probability distributions are used to model
states of uncertainty about the values of unknown quantities.
Typically, one starts with the ``prior distribution''
for unknown quantities (`parameters') $\params$, written $p(\params | \info)$
(the probability distribution for $\params$ {\em given} prior information
$\info$). This is supplemented with $p(\data | \params, \info)$,
which describes the uncertainty about the data which we {\em would} have,
if we knew the parameters (and the prior information). The
product of these is the joint prior
\begin{align}
p(\params, \data | \info) &= p(\params | \info)p(\data | \params, \info).
\end{align}
Once a specific observed dataset $\data_o$ is known, knowledge about
$\params$ is updated from the prior to the posterior
\begin{align}
p(\params | \data, \info) &\propto
    p(\params | \info)p(\data_o | \params, \info).
\end{align}
which takes the data into account.

\subsection{The shape of the model curve}
The algorithm fits the data with a model curve
$f(\x)$, defined between $\x=\x_{\rm min}$ and $\x=\x_{\rm max}$,
which is assumed to be a sum of the following components:
\begin{enumerate}
\item A background component
\item $N_{\rm wide}$ wide gaussians
\item $N_{\rm narrow}$ narrow gaussians, to model the spikes.
\end{enumerate}

We compute the posterior distribution for the unknown parameters of the
components, and use \citep{dnest4} to generate samples from that posterior
distribution, representing plausible parameter values. From each sample,
it is possible to compute the crystallinity, making it straightforward to
calculate the marginal posterior distribution of the crystallinity, describing
uncertainty about its value in the light of the data.

@@

The shape of the model function $f(x)$ is therefore
\begin{align}
f(\x) &= f_{\rm bg}(\x) + f_{\rm wide}(\x) + f_{\rm narrow}(\x)
\end{align}
and the crystallinity is defined as the proportion of the area in the
wide and narrow components which is accounted for by the wide component:
\begin{align}
C &= \frac{\int f_{\rm wide}(\x) \, d\x}
          {\int f_{\rm wide}(\x) \, d\x + \int f_{\rm narrow}(\x) \, d\x}
\end{align}

>>>>

\newcommand{\hypers}{\boldsymbol{\alpha}}
\newcommand{\params}{\boldsymbol{\theta}}
\newcommand{\data}{\boldsymbol{d}}
\newcommand{\info}{I}
\newcommand{\x}{x}
\newcommand{\todo}{\color{orange} \bf}

@@

\section{Bayesian inference}
In the Bayesian framework, probability distributions are used to model
states of uncertainty about the values of unknown quantities.
Typically, one starts with the ``prior distribution''
for unknown quantities (`parameters') $\params$, written $p(\params | \info)$
(the probability distribution for $\params$ {\em given} prior information
$\info$). This is supplemented with $p(\data | \params, \info)$,
which describes the uncertainty about the data which we {\em would} have,
if we knew the parameters (and the prior information). The
product of these is the joint prior
\begin{align}
p(\params, \data | \info) &= p(\params | \info)p(\data | \params, \info).
\end{align}
Once a specific observed dataset $\data_o$ is known, knowledge about
$\params$ is updated from the prior to the posterior
\begin{align}
p(\params | \data_0, \info) &\propto
    p(\params | \info)p(\data_o | \params, \info).
\end{align}
which takes the data into account.

\section{The shape of the model curve}
The algorithm fits the data with a model curve
$f(\x)$, defined between $\x=\x_{\rm min}$ and $\x=\x_{\rm max}$,
which is assumed to be a sum of the following components:
\begin{enumerate}
\item A background component, described by parameters $\params_{\rm bg}$
\item A wide component, described by parameters $\params_{\rm wide}$
\item Narrow spikes, described by parameters $\params_{\rm narrow}$
\end{enumerate}

The total model function $f_{\rm tot}(x)$ is therefore
\begin{align}
f_{\rm tot}(\x) &= f_{\rm bg}(\x; \params_{\rm bg})
       + f_{\rm wide}(\x; \params_{\rm wide})
       + f_{\rm narrow}(\x; \params_{\rm narrow})
\end{align}
and the crystallinity is defined as the proportion of the area in the
wide and narrow components which is accounted for by the wide component:
\begin{align}
C &= \frac{\int_{\x_{\rm min}}^{\x_{\rm max}}
            f_{\rm wide}(\x; \params_{\rm wide}) \, d\x}
          {\int_{\x_{\rm min}}^{\x_{\rm max}}
            \left[f_{\rm wide}(\x; \params_{\rm wide})
            + f_{\rm narrow}(\x; \params_{\rm narrow}) \right] \, d\x}
\end{align}

We compute the posterior distribution for the unknown parameters of the
components, and use \citep{dnest4} to generate samples from that posterior
distribution, representing plausible parameter values. From each sample,
it is possible to compute the crystallinity, making it straightforward to
calculate the marginal posterior distribution of the crystallinity, describing
the remaining uncertainty about its value in the light of the data.

@@
